---
title: "Data Analysis 2"
author: "Adam Forestier, Brandon Jones, Nathaniel Williams, Parika Gupta"
date: "December 1, 2017"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

# Introduction

The primary goal of our project is to determine if we can closely predict movie ratings before critic ratings are released and the films are given an average user score on IMDb. Being able to determine whether or not a movie will be a quality release without critic ratings could address a number of business problems that surround the financing and marketing strategies for different types of movies. Film studios and movie theatres could use these insights to budget accordingly and properly advertise movies to maximize popularity, ratings, and profits. The dataset we will be using contains 5,043 movies, and 28 variables related to the movie.

The primary research questions we wish to ask revolve around predicting a movie's IMDb score. One of which is will the number of faces in a movie poster correlate with the movie's rating? Can the multiple actor variables and the director variable be useful in predicting a movie's IMDb score? Does a movie's parental guidance rating (meaning R, PG-13, etc.) play into how well a movie scores? Does a movie's budget help determine a movie's critical success? Do certain genres tend to receive higher critic scores?

The secondary questions we are interested in solving revolve around being able to determine a movie's revenue based on the following criteria: Do different genres of movie play into a film's revenue total? Do critic scores determine how much a movie grosses? Does the number of faces on a poster play into how much a movie grosses? How much do popular actors and directors determine a movie's financial success?

Tertiary questions we wish to investigate are general questions about how variables relate to each other. We wish to find out if different nation's films tend to have different duration times. We would like to see how gross compares to budget and determine what budget expenditure typically leads to the best returns ratio-wise. We also would like to determine what key plot words are used most commonly, and if they lead to commercial and critical success.

\newpage

# Data Source and Collection

Our data was retrieved as a CSV file from Kaggle:

https://www.kaggle.com/deepmatrix/imdb-5000-movie-dataset

The data was retrieved by scraping 28 values for 5043 movies from the IMDb website and by using facial recognition software run on 4906 posters. The movies' release dates span over 100 years and the movies originate from 66 different countries. The variables of the dataset are as follows:

* movie_title = Name of the movie
* color = Whether the movie was shot in color or black and white
* num_critic_for_reviews = How many critics reviewed the movie
* movie_facebook_likes = How many likes the movie gained on facebook
* duration = How long the movie lasts
* director_name = Name of director
* director_facebook_likes = How many facebook likes the director has received 
* actor_3_name = Actor's name
* actor_3_facebook_likes = How many facebook likes the actor has received 
* actor_2_name = Actor's name
* actor_2_facebook_likes = How many facebook likes the actor has received
* actor_1_name = Actor's name
* actor_1_facebook_likes = How many facebook likes the actor has received
* gross = How much money the movie generated while in theatre
* genres = Type of movie
* num_voted_users = Number of critics who gave the movie a score on IMDB
* cast_total_facebook_likes = Number of facebook likes the cast as a whole received
* facenumber_in_poster = How many actor's faces appear in the 
* plot_keywords = Keywords users gave for the movies plot
* movie_imdb_link = link to movie on IMDB Page
* num_user_for_reviews = Number of people who gave the movie a score on IMDB
* language = Original language of movie
* country = Country movie was made in
* content_rating = The content rating of the movie
* budget = Amount of money cast and crew had to make movie
* title_year = Year movie was released
* imdb_score = Average scire the film received
* aspect_ratio = the ratio of the width to the height of a screen

Some of our data will need to be cleansed before moving on to analysis. The variable "director_facebook_likes" is inaccurate in many instances. Foreign films using currencies different from USD may have a misrepresented gross revenue, because the values aren't equivalent with foreign currencies. This can be addressed by transforming these values by an exchange rate. We would also like to retrieve the release date for each film, since that variable was not included in the original dataset.
```{r}

```

The goal of our project is to find connections between variables and the film's IMDb score. We believe that the key variables are the following: facenumber_in_poster, director_name, actor_1_name, actor_2_name, and actor_3_name. Our hypothesis is that these variables will be crucial in predicting a movie's overall critic score. We also hypothesize that budget, gross, actors, directors, genre and content rating will play into how much a movie grosses in revenue.


# Preperation work 
Clear the environment 
```{r}
rm(list = ls())
```

Load piping Package
```{r}
# Note: User may need install packages if they have not previously used them
library(magrittr)
```

Load Data Manipulation Packages
```{r}
library(plyr)
library(dplyr)
```

# Read in data 

Set Working Directory
```{r}
setwd("D:/College Work/Junior Year/IST 3420/Project/Files")
```

Load movie CSV file in as a data frame  
```{r}
movie_df <- read.csv("movie_metadata CSV.csv", header = TRUE, stringsAsFactors = F) 

# Change blank spaces to NA
movie_df[movie_df==""] <- NA
```


show structure of movie_df
```{r}
print(str(movie_df))
```

\newpage

# Data cleansing and manipulation

Most of the variables for the IMDb data set were formatted correctly to begin with, but some cleansing was required. 

See if there are any observations with NA values for title or score making analysis useles
```{r}
count(movie_df[!(is.na(movie_df$movie_title)) & !(is.na(movie_df$imdb_score)), ])
```

It appears that every observation has a score and title


Convert categorical variables into factors to improve analyses

Color is a factor but is currently a character type, need to change it to a number. 1 = "Color", 0 = "Black and White"
```{r}
movie_df$color <- ifelse(movie_df$color == "Color", 1, 0)
```


Convert categorical to factors
```{r}
# Create list of categorical variable columns that are to be converted to factors
factor_cols <- c("color", "facenumber_in_poster", "language", "country", "content_rating", "title_year", "aspect_ratio")

# Use lapply to convert factor_cols to factor types in movie_df
movie_df[factor_cols] <- lapply(movie_df[factor_cols], factor)

# Show structure of converted variables
print(str(movie_df[factor_cols]))
```

We created a new variable that calculated the gross profit of a film by subtracting the budget of the film from the gross revenue. This will allow us to perform analyses later on and see how other factors may affect the profit of movies.

create new variable gross_profit by subtracting budget from gross
```{r}
# Create new vector, positive means budget was lower than gross, negative means that budget exceeds gross
movie_df$gross_profit <- movie_df$gross - movie_df$budget
```


The values of the "movie_title" variable in our data set had the character "Â" appended at the end and white space at the end, so we removed the character and trimmed the white space for better readability and to prevent any issues that might have arisen without eliminating the extra white space. 
```{r}
# Remove the Â at the end of the titles
movie_df$movie_title <- gsub("Â", "", movie_df$movie_title) 

# trim whitespace
movie_df$movie_title <- trimws(movie_df$movie_title)
```

There were also a number of duplicated values of movies, so we checked for redundant data and removed duplicate entries.
```{r}
# Show number of unique titles
length(unique(movie_df$movie_title))

# Show number of duplicated titles
anyDuplicated(movie_df$movie_title)
```

After verifying these are actually duplicates, We need to remove duplicates to avoid skew in our findings
```{r}
# This statement removes movie titles that are duplicates
# movie_df<- movie_df[!duplicated(movie_df$movie_title movie_df$director_name && movie_df$title_year),]

#dup_movies <- 
```

Show summary statistics now that the dataset has been cleansed

Load stargazer to improve appearance of summary statistics
```{r}
library(stargazer)
```


```{r}
stargazer(movie_df, type = "text", title = "Movie Data frame Summary Statistics")

# Stargazer only shows statistics for non categorical and character types
```

\newpage

# Data Visualization and analysis

Load plotly visualization package
```{r}
# Note: plotly package graphs are interactive
library(plotly) 
library(ggplot2)
```


We can look to see which years tend to have higher quality movies based on average IMDb scores of movies over time.

```{r}
# Create data frame with year and average score
yearly_score <- ddply(movie_df, c("title_year"), summarise,
                      mean(imdb_score, na.rm = TRUE))
colnames(yearly_score) <- c("title_year", "avg_score")

# Plotly line graph that shows average imdb scores across the years
plot_ly(yearly_score, x = yearly_score$title_year, y = yearly_score$avg_score, type = "scatter", mode = "lines") %>% 
  layout(title = "Year and Score") 
```

Up until the 1970s, only a few movies may be registered in the IMDb database for a given year, with many years only yielding a single movie. This leads to large variances between years up until the 1970s, such as how a single movie in 1920 is completely representative of the year's extremely low average movie IMDb score of 4.8, only to surge the exisitng data point because a single movie performed better in 1925 and received a rating of 8.3. Regardless, the general trend appears to be a declining average IMDb score over time. This may not reflect the quality of large production films, but rather reduced barriers to entry have enabled a surge in indie films to enter the market and the fact that the IMDb database can also register these films with the rise of internet-based titles being published on platforms like YouTube. 


Show number of movies released in each year

Create df with number year and movies in that year
```{r}
movie_yearly_count <- movie_df %>% 
  group_by(title_year) %>%
  summarise(yearly_total = sum(length(title_year)))
```

Verify sum of number of movies in 
```{r}
sum(movie_yearly_count$yearly_total)
```

Plotly line graph that shows number of films released by year in the data set
```{r}
plot_ly(movie_yearly_count, 
        x = movie_yearly_count$title_year, y = movie_yearly_count$yearly_total, type = "scatter", mode = "lines") %>% 
  layout(title = "Total films released by year")

# Plot skips some years along the x-axis, but this was for spacing; the interacting with the chart will show count for every year
```

Load ggplot2 package
```{r}
library(ggplot2)
```

We can explore which actors have the highest average movie score overall.
```{r}
# Find actors with highest average movie score 
actor_rating_df <- ddply(movie_df, c("actor_1_name"), summarise, # Pull out actor 1 and rating data
                      actor_m <- mean(imdb_score, na.rm = TRUE),  # Find the mean of actors and discount na values
                      number <- length(na.omit(imdb_score))) # Count number of movies, but omit na values
colnames(actor_rating_df) <- c("actor_name", "avg_score", "num_of_movies") # Rename headers
actor_avg <- actor_rating_df[which(actor_rating_df$num_of_movies >= 7),]# Actors in at least 7 movies to avoid skew
actor_avg <- actor_avg[which(actor_avg$avg_score >= 7.15),] # Show only highest scoring actors for readability

# Convert actor_name to a factor
actor_avg$actor_name <- factor(actor_avg$actor_name)

# Show highest to lowest average score
actor_avg$actor_name <- reorder(actor_avg$actor_name, actor_avg$avg_score)

# Visualization of highest scoring actors
ggplot(actor_avg, aes(x = avg_score, y = actor_name)) + # Sets x to average score and y to the name
  geom_point(aes(colour = actor_name)) + # Makes a point graph with different colors per actor
  theme(axis.text = element_text(size = 10)) + # Sets font size
  ggtitle("Actor Scores") + # Changes chart title
  xlab("Average Score") + # Changes x field title
  ylab("Actor Name")  # Changes y field table
```

We can see a visual representation of actors with the highest average IMDb ratings for the films they star in, with the top 3 being Leonardo DiCaprio, Tom Hanks, and Clint Eastwood.


We can explore which directors have the highest average movie score overall.
```{r}
# Find actors with highest average movie score 
dir_rating_df <- ddply(movie_df, c("director_name"), summarise, # Pull out director name and rating data
                      dir_m <- mean(imdb_score, na.rm = TRUE),  # Find the mean of directors and discount na values
                      number <- length(na.omit(imdb_score))) # Count number of movies, but omit na values
colnames(dir_rating_df) <- c("director_name", "avg_score", "num_of_movies") # Rename headers
dir_avg <- dir_rating_df[which(dir_rating_df$num_of_movies >= 7),]# Direct  at least 7 movies to avoid skew
dir_avg <- dir_avg[which(dir_avg$avg_score >= 7.5),] # Show only highest scoring directors for readability
dir_avg <- dir_avg[which(dir_avg$director_name != ''),] # Exclude movies who have no director name 

# Convert actor_name to a factor
dir_avg$director_name <- factor(dir_avg$director_name)

# Show highest to lowest average score
dir_avg$director_name <- reorder(dir_avg$director_name, dir_avg$avg_score)

# Visualization of highest scoring actors
ggplot(dir_avg, aes(x = avg_score, y = director_name)) + # Sets x to average score and y to the name
  geom_point(aes(colour = director_name)) + # Makes a point graph with different colors per director
  theme(axis.text = element_text(size = 10, colour = 'black')) + # Sets font size
  ggtitle("Director Scores") +  # Changes chart title
  xlab("Average Score") + # Changes x field title
  ylab("Director Name") # Changes y field table
```

We can see a visual representation of directors with the highest average IMDb ratings for the films they star in, with the top 3 being Christopher Nolan, Quentin Tarantino, and James Cameron.


We can explore the average IMDb score for each content rating.
```{r}
content_df <- ddply(movie_df, c("content_rating"), summarise, # New DF w/ rating, avg score and # of movies 
                    mean(imdb_score, na.rm = TRUE), # Creates column containing mean scores
                    length(na.omit(imdb_score))) # Omits scores with NA values
colnames(content_df) <- c("content_rating", "avg_score", "number_of_films") # Rename columns
content_df <- content_df[which(content_df$content_rating != ""),] # Eliminate films from analysis without rating

# Visualization of average score by content rating
ggplot(content_df, aes(x = content_rating, y = avg_score)) + # Bar graph with x as rating and y as score
  geom_bar(stat = "identity", width = .5, fill = 'deepskyblue3') + # Creates space between bars and changes color
  theme(axis.text = element_text(size = 8, colour = 'black', angle = 45)) + # Sets text size to 8, text color to black, and angles text for separation purposes
  ggtitle("Content Scores") + # Title of graph
  xlab("Content Rating") + # Title of x axis
  ylab("IMDB Score") # Title of y axis
```

Based on this plot, we can see the highest performing category of movies based on content rating is TV-MA (intended for a mature audience), with the lowest performing category being PG-13.This could possibly be the effect of restricted possibilities of movies that target a PG-13 audience not being as favored by mature audiences who watch both TV-MA movies and PG-13 movies. 


We can show the spread of scores per content rating category with box plots to detect if outliers could be affecting the score averages of each content rating category.
```{r}
# Create data frame that excludes movies with no content rating 
content_df2 <- movie_df[which(movie_df$content_rating != ''), ]

#Use plotly for great boxplot capabilities, note: Might need to install "plotly"
plot_ly(content_df2, x = content_df2$imdb_score, color = content_df2$content_rating, type = "box") %>%
  layout(title = "Movie Type and Score Range")
```

Based on these box plots, we can see a number of potential outliers in several categories. There also appear to be clusters of values picked up as outliers in the content rating categories of R, PG-13, and PG, which may indicate bimodal distributions in these categories. These outliers will need to be accounted for when considering the content rating's effect on the average movie score.


Explore Average IMDB rating by number of faces in poster
```{r}
# Create df with face numbers in poster and respective average score
face_df <- movie_df %>%
  group_by(face_count = facenumber_in_poster) %>%
  summarise(avg_score = mean(imdb_score)) %>%
  na.omit() # Omit NA values

# Bar Graph of data
ggplot(face_df, aes(x = face_count, y = avg_score)) +
  geom_bar(stat = "identity", width = .5, fill = "deepskyblue3") +
  xlab("Number of Faces") +
  ylab("Average Score") +
  ggtitle("Average Score by faces in poster") +
  theme(axis.text = element_text(size = 8, angle = 45))
  
```

Show 5 number summary for average score based on color using boxplot
```{r}
# Create Data Frame with average score and color
color_film <- movie_df %>%
  select(color, imdb_score) %>% na.omit()

ggplot(color_film, aes(x = color, y = imdb_score, fill = color)) +
  geom_boxplot() +
  xlab("Color Type") +
  ylab("Score") +
  ggtitle("Black and White vs Color Film Scores")
```

Visualize how language affects movie score
```{r}
language_df <- movie_df %>%
  group_by(language) %>%
  summarise(avg_score = mean(imdb_score)) %>% na.omit()

ggplot(language_df, aes(x = language, y = avg_score)) +
  geom_bar(stat = "Identity", width = .5, fill = "deepskyblue3") +
  theme(axis.text = element_text(size = 10, angle = 90)) + 
  xlab("Original Language") +
  ylab("Average Score") +
  ggtitle("Language and Score")
```

Visualize how country the film shot in affects average score
```{r}
country_df <- movie_df %>%
  group_by(country) %>%
  summarise(avg_score = mean(imdb_score)) %>% na.omit()

ggplot(country_df, aes(x = country, y = avg_score)) +
  geom_bar(stat = "Identity", width = .5, fill = "deepskyblue3") +
  theme(axis.text = element_text(size = 8, angle = 90))
```

\newpage

# Correlation

We chose to use Pearson Correlation, because the measures are not ordinal, and measures are linear

Pearson Correlation Table 1

Load Corrplot
```{r}
library(corrplot)
```

Clear out global environment except for movie_df for information management purposes
```{r}
rm(list = ls()[!ls() %in% c("movie_df")])
```

```{r}
# Create data frame with only numeric types to see how numeric variables correlate to imdb score 
numeric_df <- select_if(movie_df, is.numeric) # Other option cor(movie_df[sapply(movie_df, is.numeric)])

# Creates a correlation table showing variables correlate w/ each other by number and text color 
num_correlation <- round(cor(numeric_df, use = "pairwise.complete.obs", method = "pearson"),1)
corrplot(num_correlation, method = "number", type = )

# Note: After running this chunk, you can also select hte num_correlation item from the global environment to view a generic table
```

Pearson Correlation Table 2

```{r}
# Creates a correlation table showing variables correlate w/ each other by color and  circle size
corrplot(num_correlation, method = "circle")
```

Clear out global environment except for movie_df for information management purposes
```{r}
rm(list = ls()[!ls() %in% c("movie_df")])
```




