---
title: "Simple Linear Regression"
author: "Langtao Chen"
date: "Oct 27, 2017"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, Message = F)
```

&nbsp;
&nbsp;


\newpage


```{r}
# Clean the environment
rm(list = ls())
```


# 1. Read in Data

In this example, we'll use the shoesize dataset to demonstrate how to conduct simple regression analysis.

```{r}
df <- read.csv("shoesize.csv")
```

```{r}
head(df)
```

```{r}
# Remove the index column
df$Index <- NULL
str(df)
```

# 2. Explore Data

Before conducting regression analysis, we can draw a scatterplot matrix to explore the relationships between variables.

```{r}
# plot(df)
```

From the above scatterplot matrix, we see a clear positive linear relationship between shoe size and the person's height. Let's zoom in this relationship.

```{r}
plot(x = df$Height, y = df$Size, col = "red",
     xlab = "Height", ylab = "Shoe Size")
```

# 3. Simple Linear Regression

With the visualization of linear relationship between shoe size and height, now let's formally use linear regression model to analyze this linear relationship.

```{r}
# Regress shoe size on height
model1 <- lm(Size ~ Height, data = df)

# Show the result
summary(model1)
```

Though the summary() provides us everything about the linear regression analysis, the output is not professional and not easy to read.

Let's use the stargazer package to report regression results in a more professional way.

```{r}
# install.packages("stargazer") #Install stargazer package, do this only once
library(stargazer)
```


```{r}
stargazer(model1, type = "text", star.cutoffs = c(0.05, 0.01, 0.001),
          title="Simple Linear Regression", digits=4)
```


The effect of height on shoe size is significant (p-value < 0.05). If height increases by one inch, shoe size would increase by 0.43 unit on average. 75.85% of the variance of shoe size can be explained by height.

From the regression result, we can build the following relationship between shoe size and height:

ShoeSize = -19.3266 + 0.4273*Height

We can use this formula to predict shoe size if a person's height is provided. For example, if Tommy is 79 inches in height, then the shoe size is predicted as:

```{r}
cat("Shoe Size = ", -19.3266 + 0.4273*79)
```

