---
title: "Homework 8"
author: "Lorenzo Lewis"
date: "10/16/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear the global environment
```{r}
rm(list=ls())
```

Load in required libraries
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

# 2
Read the data file into R
```{r}
df <- read.csv("mst_enrollment2015.csv")
```

# 3
Use dplyr::filter() to remove all non-US data.
```{r}
mst <- dplyr::filter(df, Country == "United States")
```

# 4
Use dplyr::filter() to remove data of Missouri (in-state enrollment)
```{r}
mst <- dplyr::filter(df, State != "Missouri")
```

# 5
Use tidyr package to reshape the dataset by transforming it from wide format to long format, using "Year" as key and "Enroll" as value. 
```{r}
mst <- gather(df, Year, Enroll, X2011, X2012, X2013, X2014, X2015)
```

# 6
Use dplyr package to aggregate enrollment data into state level, i.e. calculate total enrollment for each state. 
```{r}
mst <- group_by(df, State)
mst <- summarize(df, pop = sum(Enroll))
```

#8
Use ggplot2 package to draw the MST out-of-state enrollment plot on the US states map.

Prepare the dataset
```{r}
us_states <- map_data("state")
colnames(df)[1] <- "region"
df$region <- tolower(df$region)
df <- df[(df$region %in% us_states$region),]
us_states <- left_join(df, us_states, by = "region")
```

Plot the dataset
```{r}
ggplot(data = us_states) +
  geom_polygon(aes(x = long, y = lat, fill = pop, group = group), color = "white") +
  coord_fixed(1.3) +
  labs(title = "Missouri S&T Out of State Enrollment",
       fill = "Enrollment",
       x = "",
       y = "") +
  scale_x_continuous(breaks = c()) +
  scale_y_continuous(breaks = c())
```


