---
title: "Homework 8"
author: "Adam Forestier"
date: "October 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

Clean Environment
```{r}
rm(list = ls())
```

Load Packages
```{r}
require(dplyr)
require(magrittr)
require(ggplot2)
require(tidyr)
require(maps)
```

# Read MST Enrollment data 
```{r}
mst <- read.csv("mst_enrollment2015.csv")

# Show structure
str(mst)
```

# Remove not US Data
```{r}
mst_us <- mst %>% dplyr::filter(mst$Country == "United States")
```

# Remove Missouri Data
```{r}
mst_us <- mst_us %>% dplyr::filter(mst_us$State != "Missouri")
```

Load State Map Data
```{r}
# Create a data frame of map data for all states
us_states <- map_data("state")

# List all states in map dataset
map_state <- unique(us_states$region)
map_state
```

We see that Armed Forces Europe, Armed Forces Pacific, Hawaii, and Alaska aren't in the map package so we need to remove them 
```{r}
mst_us <- mst_us %>% dplyr::filter(mst_us$State != "Armed Forces Europe")
mst_us <- mst_us %>% dplyr::filter(mst_us$State != "Armed Forces Pacific")
mst_us <- mst_us %>% dplyr::filter(mst_us$State != "Alaska")
mst_us <- mst_us %>% dplyr::filter(mst_us$State != "Hawaii")
```


# Reshape the dataset by transforming it from wide format to long format
```{r}
mst_us_long <- mst_us %>% gather(Year, Enroll, X2011, X2012, X2013, X2014, X2015)
```

lower capitilization on State names
```{r}
mst_us_long$State <- tolower(mst_us_long$State)
```

# Aggregate enrollment data into state level, i.e. calculate total enrollment for each state using dplyr
```{r}
# Aggregate data
by_state <- group_by(mst_us_long, State)
enrollment_count <- summarise(by_state, pop = sum(Enroll))

# Show top 6 enrollment states 
head(enrollment_count[order(enrollment_count$pop, decreasing = T),])
```

# MST out-of-state enrollment plot on the US states map

Merge Data Sets
```{r}
# Rename Columns of us_states for merging purposes
colnames(us_states) <- c("long", "lat", "group", "order", "State", "subregion")

# Merge Enrollment_Count and us_states for map
pop_map <- merge(us_states, enrollment_count, by="State")
```


Create Map Visualization
```{r}
ggplot() + 
  geom_polygon(data = pop_map,
               aes(x=long, y=lat, group = group, fill=pop_map$pop),
               colour="black") + 
  scale_fill_continuous(low = "darkslategray1", high = "darkblue", guide="colorbar") + 
  theme_bw() +
  labs(fill = "Enrollment",
       title = "Missouri S&T Enrollment by State",
       x="", y="") +
  scale_y_continuous(breaks=c()) +
  scale_x_continuous(breaks=c()) + 
  theme(panel.border =  element_blank())

```
