---
title: "Homework 7"
author: "Adam Forestier"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

Clear the Environment
```{r}
rm(list = ls())
```

Load in packages
```{r}
require(jsonlite)
require(dplyr)
```

set directory
```{r}
setwd("D:/College Work/Junior Year/IST 3420/Homework - Code/Homework 7")
```


# Read in JSON File Data

```{r}
# Read in data
json_df <- stream_in(file("weblog_clean.jsonl"))
```

# Change request.datetime as POSIXct  
```{r}
json_df <- json_df %>% mutate(request.datetime = as.POSIXct(request.datetime))
```

# Show summary statistics
```{r}
summary(json_df)
```


# Calculate Frequency distribution of weekday
```{r}
# Calculate frequency distribution of weekday and put it in column format ordered greatest to least
day_freq <- cbind(table(json_df$weekday))
day_freq <- cbind(day_freq[order(-day_freq),])
print(day_freq)
```

# Crosstabulate weekday and request.uri. 
```{r}
cross_tbl <- xtabs(~ request.uri + weekday, data=json_df)
print(cross_tbl)
```

# Bar plot to show the distribution of visits by weekday
```{r}
# Create table
weekday_distribution <- table(json_df$weekday)

# Bar graph
barplot(weekday_distribution,
        las = 2,
        main="Distribution of Visits by Weekday",
        col=rainbow(7))
```

# Stacked bar plot to show the distribution of visits by request.uri and request.method 
```{r}
# Create Table 
request_info <- table(json_df$request.uri, json_df$request.method)

# Stacked Bar
barplot(request_info,
        main="Distribution of Visits by request.uri and request method",
        xlab = "Request Method",
        col = rainbow(6),
        ylim = c(0, 140000),
        legend = rownames(request_info))
```


# Pie chart to present the distribution of visits by request.uri by count
```{r}
# Create Table
uri_request <- table(json_df$request.uri)

# Create Label to include on pie chart
pie1_lbl <- paste(names(uri_request), " \n", uri_request, sep="")

pie(uri_request,
    labels = pie1_lbl,
    main = "Pie Chart with URI Visitor Count",
    col = rainbow(length(pie1_lbl)))
```

# Pie chart to present the distribution of visits by request.uri by perecentage
```{r}
# Create percent and label
pct <- round(uri_request/sum(uri_request)*100, digits = 1)
pie2_lbl <- paste(names(uri_request), " \n", paste(pct,"%"), sep="")

pie(uri_request,
    labels = pie2_lbl,
    main = "Pie Chart with URI Percentage of Visitors",
    col = rainbow(length(uri_request)))
```

# Five-number summary for "response.size". 
```{r}
# Create df with the response size column going from smallest to largest
ordered_size_df <- json_df[order(json_df$response.size),]

quantile(ordered_size_df$response.size, type = 7)
```















