---
title: "Homework 11"
author: "Adam Forestier"
date: "November 12, 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```



\newpage 


Clean current environment
```{r}
rm(list = ls())
```

Load piping and manipulation packages
```{r}
library(dplyr)
library(magrittr)
```


# Read in CSV data
```{r}
telecom_df <- read.csv("Telco-Customer-Churn.csv", header = T)

# Remove missing data
# telecom_df <- na.omit(telecom_df)
```


# Transform "Churn" to numeric type

Show current class of churn
```{r}
# See current type of Churn
class(telecom_df$Churn)
```

Show levels of churn
```{r}
levels(telecom_df$Churn)
```

Transform churn to numeric
```{r}
# Create Churn column with 1 as yes and 0 as no
telecom_df$Churn <- ifelse(telecom_df$Churn=="Yes", 1, 0)

class(telecom_df$Churn)
```


# Show summary statistics of the dataset
```{r}
summary(telecom_df)
```


# Calculate the percentage of customers who churned
```{r}
# Take length of 1 divided by total number of observations multiply by 100 and round decimal to 1 
round(length(which(telecom_df$Churn == 1))/length(telecom_df$Churn) * 100, 1)
```


# Multiple Linear Regression model to fit the data, regressing on churn

Using stepAIC() method in MASS package to determine which independent variables to use

```{r}
# Load Package
library(MASS)
```

Regress Churn on all other variables and use step method to determine which model to use
```{r}
# Regress mpg on all other variables
fit <- lm(Churn ~ ., data = telecom_df)

# Perform stepwise model selection
step <- stepAIC(fit, direction = "both")

# Show the results
step$anova
```

According to the stepwise variable selection approach, the recommended model is SeniorCitizen + Dependents + Tenure + MultipleLines + InternetService + OnlineSecurity + OnlineBackup + TechSupport + StreamingTV + StreamingMovies + Contract + PaperlessBilling + PaymentMethod + TotalCharges


Use model selected by stepwise selection as the final model
```{r}
# Regress churn on other variables
# telecom_df <- na.omit(telecom_df)
final_linear <- lm(Churn ~ SeniorCitizen + Dependents + Tenure + MultipleLines + InternetService + OnlineSecurity + OnlineBackup + TechSupport + StreamingTV + StreamingMovies + Contract + PaperlessBilling + PaymentMethod + TotalCharges,
                data = telecom_df)



# Show summary of final fit regression model
summary(final_linear)
```

Detach MASS package to regain use of the select() method in dplyr package
```{r}
detach("package:MASS", unload = F)
```

# Logistic regression model to fit data, regressing on Churn
```{r}
# Regress churn on other variables
final_logistic <- glm(Churn ~ SeniorCitizen + Dependents + Tenure + MultipleLines + InternetService + OnlineSecurity + OnlineBackup + TechSupport + StreamingTV + StreamingMovies + Contract + PaperlessBilling + PaymentMethod + TotalCharges, family=binomial(link='logit'),
                data = telecom_df)

# Show the summary of logistic regression results
summary(final_logistic)
```

# Stargazer() function to show results of both linear and logistic regression analyses

Load Package
```{r}
library(stargazer)
```

Show side-by-side comparison using stargazer
```{r}
stargazer(final_linear, final_logistic, type = "text", star.cutoffs = c(0.05, 0.01, 0.001),
          title="Multiple Linear vs. Logistic Regression", digits=4)
```

Show why there are only 7,032 observations instead of 7,043
```{r}
sum(is.na(telecom_df))
```

There are 11 na values in the data frame

# Calculate the McFadden R2 for the logistic regression model

create null model
```{r}
# Fit null model
null_model <- glm(Churn ~ 1,family=binomial(link='logit'),data=telecom_df)

# Show likelihood of null model
logLik(null_model)
```

Calculate McFadden R2
```{r}
cat("McFadden pseudo R2 = ", 1-logLik(final_logistic)/logLik(null_model))
```





