---
title: "Homework 12"
author: "Adam Forestier"
date: "November 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```


Clear Current Environment
```{r}
rm(list = ls())
```


Load Piping package
```{r}
library(magrittr)
```

# Read CSV File data 
```{r}
bike_df <- read.csv("DC_Bike_Rentals.csv", header = T)

# Remove na values
bike_df <- na.omit(bike_df)

# Explore data
head(bike_df)
```


# Show the structure of the data
```{r}
print(str(bike_df))
```


# Transform hour, season, holiday, workingday, & weather into factors
```{r}
# Choose variables to coerce to factors 
factor_cols <- c("hour", "season", "holiday", "workingday", "weather")

# Convert variables to factors
bike_df[factor_cols] <- lapply(bike_df[factor_cols], factor)

print(str(bike_df))
```
 
# Show summary statistics of the data set

Load stargazer package to improve appearance of output
```{r}
library(stargazer)
```
 
Show summary statistics output 
```{r}
stargazer(bike_df, type = "text", title = "DC Bike Rentals")
```
 
# Create 10-fold validation configuration of control

Load caret package 
```{r}
library(caret)
```

Cross-validation
```{r}
fit_control <- trainControl(method = "cv", number = 10) 
```
 
# Train multiple linear regression model to predict count
```{r}
# Train linear model
set.seed(1234)
multi_fit <- train(count ~ ., data = bike_df,
                   trControl = fit_control,
                   method="lm")

# Show results
print(multi_fit)
```
 
 
# Train GBM model that predicts count 
```{r}
# Train GBM model
set.seed(1234)
GBM_fit <- train(count ~ ., data = bike_df,
                 trControl = fit_control,
                 method = "gbm", verbose = F)

print(GBM_fit)
```
 
# Compare the two different predictive models

Resample and show results
```{r}
results <- resamples(list(linear = multi_fit, GBM = GBM_fit))

summary(results)
```
 
 
 
 
 