---
title: "Homework 5"
author: "Adam Forestier"
date: "September 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Clear the current environement
```{r}
rm(list = ls())
```

# 1 Download the CSV file

# Read the CSV into R
```{r}
# set working directory
setwd("D:/College Work/Junior Year/IST 3420/Homework - Code/Homework 5")

# Read in CSV File
retailer_info <- read.csv("top100retailers2015.csv", header = TRUE)
```

# 3 Show summary statistics of the dataset
```{r}
summary(retailer_info)
```

# 4 Convert factor data into numeric data for all columns except "Company" and "Headquarter".
```{r}
retailer_info$X <- as.numeric(retailer_info$X)
retailer_info$Rank <- as.numeric(retailer_info$Rank)
retailer_info$RetailSales2014 <- as.numeric(gsub('[$,]', '', retailer_info$RetailSales2014))
retailer_info$USASalesGrowth <- as.numeric(gsub('[%]', '', retailer_info$USASalesGrowth))
retailer_info$WorldwideRetailSales <- as.numeric(gsub('[$,]', '', retailer_info$WorldwideRetailSales))
retailer_info$USAPercentageOfWorldwideSales <- as.numeric(gsub('[%]', '', retailer_info$USAPercentageOfWorldwideSales))
retailer_info$Stores2014 <- as.numeric(gsub('[,]', '', retailer_info$Stores2014))
retailer_info$Growth <- as.numeric(gsub('[%]', '', retailer_info$Growth))

```


# 5 Create a column named "City" and use it to store the city extracted from "Headquarter"
```{r}
city <- gsub(",.*$", "", retailer_info$Headquarter)
retailer_info <- cbind(retailer_info, city)
```

# 6 Create a column named "State" and use it to store the state extracted from "Headquarter"
```{r}
state <- gsub("^.*, ", "", retailer_info$Headquarter)
retailer_info <- cbind(retailer_info, state)
```

# 7 Show summary statistics of the new dataset
```{r}
summary(retailer_info)
```

