---
title: "Homework 6"
author: "Adam Forestier"
date: "October 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clean the current environment 
```{r}
rm(list = ls())
```

# Read in dataset

```{r}
apache_log <- read.table("sample_apache.log", sep = "", stringsAsFactors = F)
```

# Show the raw data

Show the structure of data
```{r}
str(apache_log)
```

Show first 10 values
```{r}
head(apache_log, n = 10)
```

# Transform and cleanse the data

Load in dplyr package
```{r}
library(dplyr)
```

Create the new variables
```{r}
apache_log <- apache_log %>% mutate(remote.host = V1,
                                    status = V7,
                                    request.datetime = as.POSIXct(V4, format = "[%d/%b/%Y:%H:%M:%S"),
                                    request.date = as.Date(request.datetime),
                                    weekday = weekdays(request.datetime),
                                    request.method = gsub(" ","",substring(V6,1,4)),
                                    request.uri = gsub("/","",substring(V6,6,nchar(V6)-9)))
                                    
```

Remove the unnecessary variables
```{r}
# Selects only the six variables needed
apache_log <- apache_log %>% select(remote.host, status, request.datetime, weekday,request.method, request.uri)
```

# Show visitors per ip address

Create data frame which shows the total number of visits for each unique IP address
```{r}
visit_per_IP <- count(apache_log, remote.host)
colnames(visit_per_IP) <- c("remote.host", "count_of_visit")
```

Show total count of each number of visits per unique IP 
```{r}
table(visit_per_IP$count_of_visit)
```


Subset the "visit_per_IP" to create a data frame called "visit_per_IP2" which only contains information for unique IP addresses that have 2 or more visits. 
```{r}
visit_per_IP2 <- visit_per_IP[which(visit_per_IP$count_of_visit >= 2),]
```





