---
title: "Homework 10"
author: "Adam Forestier"
date: "November 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, Message = F)
```


\newpage


# Task 1

Clean Current environment
```{r}
rm(list = ls())
```
Load dplyr and magrittr packages
```{r}
library(magrittr)
library(dplyr)
```

# 1 Use mtcars dataset and focus on miles per gallon and transmission
```{r}
task1_df <- mtcars %>% select(mpg, am)
print(head(task1_df))
```

# 2. Side by side boxplot showing the difference of mpg by transmission type
```{r}
# Create side by side boxplot using base R
two_box <- boxplot(mpg ~ am, data = task1_df, col = c("blue", "red"),
                   main = "MPG by Transmission Type",
                   xlab = "Transmission",
                   ylab = "MPG")

# Print Boxplot
print(two_box)
```

# 3 Use unpaired two sample t-test to compare the two different transmission types
```{r}
transmission1 <- task1_df$mpg[task1_df$am == 0]
transmission2 <- task1_df$mpg[task1_df$am == 1]

t.test(transmission1, transmission2, alternative = "two.sided", mu = 0, paired = F)
```


\newpage 

# Task 2


# 1 Use mtcars dataset and focus on miles per gallon and weight
```{r}
task2_df <- mtcars %>% dplyr::select(mpg, wt)
print(head(task2_df))
```


# 2 Regress mpg on wt using mpg as dependent variable, wt as independent variable. 

Visualize the data
```{r}
regression_plot <- plot(x = task2_df$wt, y = task2_df$mpg, col = "red", main = "MPG relationship with Weight",
     ylab = "MPG", xlab = "Weight")
regression_plot <- abline(lm(task2_df$mpg ~ task2_df$wt), col = "black")
print(regression_plot)
```


Regression results
```{r}
first_linear <- lm(mpg ~ wt, data = task2_df)

summary(first_linear)
```

# 3 Stargazer() function to beautify regression output

Load stargzer library
```{r}
library(stargazer)
```

Improve regression output appearance with stargazer
```{r}
stargazer(first_linear, type = "text", star.cutoffs = c(0.05, 0.01, 0.001), title = "MPG and weight simple linear regression", digits = 4)
```

# 4 Find regression model

```{r}
# Regression model = 37.2851 - 5.3445(wt/1000)

# Predicted mpg of car weighing 5500 pounds
cat("Car mpg = ", 37.2851 - 5.3445*(5500/1000))
```












